Neighbor-Based Decentralized Training Strategies for Multi-Agent Reinforcement Learning

### 2024-10-13 paper-2024-sac-marl-neighbouring-strategies.tex
final pass: improve some aspects of the paper

\begin{abstract}
Multi-agent deep reinforcement learning (MADRL) is rapidly emerging as a powerful paradigm for training intelligent agents in complex, multi-agent environments.
%
Extending beyond single-agent reinforcement learning, MADRL enables the development  of both collaborative and competitive agents in several domains, like robotics, traffic management, and video games.
%
The most widely adopted training paradigm in MADRL is centralized training with decentralized execution (CTDE), which has demonstrated significant effectiveness. 
%
However, its reliance on a centralized learner necessitates that agents acquire policies offline and subsequently execute them online. 
This inherent constraint motivates the exploration of decentralized training methodologies.
%
While these techniques offer greater flexibility, they tend to have slower convergence and can suffer from instability compared to centralized approaches.
%
Therefore, this paper investigates the effectiveness of various neighbor-based decentralized training strategies based 
  on the well-known Deep-Q Learning algorithm as a viable alternative to centralized training. 
%  
We evaluate experience sharing, k-nearest neighboring averaging, and k-nearest neighboring consensus methods 
  in a cooperative multi-agent environment and compare their performance against centralized training
  and totally decentralized training. 
%  
Our results show that neighbor-based methods can achieve comparable performance to centralized training 
  while offering improved scalability and communication efficiency. 
%
% Finally, we discuss the trade-offs between these methods and provide insights into their applicability 
%   in different scenarios.
\end{abstract}

@@

\section{Introduction}
The increasing complexity of real-world systems, from robotic swarms to smart grids, demands intelligent control strategies capable of adapting to dynamic and often unpredictable environments.
%
In this direction, \emph{reinforcement learning} (RL)~\cite{sutton2018reinforcement} has emerged as a powerful paradigm for training agents that interact with their environment and learn optimal policies through trial and error.
%
Moving beyond single-agent settings, \emph{multi-agent reinforcement learning} (MARL)~\cite{bucsoniu2010multi} extends RL to scenarios where multiple agents must collaborate or compete to achieve a common goal.
%
Recent advances in deep learning have further advanced the field, 
enabling the training of complex policies in high-dimensional state and action spaces, 
leading to the emergence of \emph{multi-agent deep reinforcement learning} (MADRL)~\cite{gronauer2022multi}.
%
MADRL has gained significant traction due to its flexibility and capacity to learn diverse behaviors, 
ranging from cooperative~\cite{gupta2017cooperative} to competitive~\cite{tampuu2017multiagent}, 
as demonstrated in various case studies involving video games~\cite{shao2019survey}, robotics~\cite{orr2023multi},
and traffic management~\cite{chu2019multi}. 
%
This paper focuses on the subdomain of \emph{cooperative multi-agent reinforcement learning}, 
a particularly active and impactful area of research within MADRL~\cite{oroojlooy2023review}.

@@

\section{Background and Motivation}\label{sec:background}
Multi-agent reinforcement learning (MARL) involves a group of \emph{agents} interacting with a shared \emph{environment}. 
%
At each timestep, agents receive \emph{observations} and select \emph{actions} according to their individual \emph{policies} (mappings from observations to actions). 
%
These individual actions combine to form a \emph{joint action}, which influences the environment's transition to a new state. 
%
Each agent then receives a scalar \emph{reward} signal, typically a function of the global state and joint action. 
%
The goal of each agent is to maximize its cumulative reward. 
%
Environments can be \emph{partially observable}, meaning agents may not have access to the full global state, only receiving local observations through their respective \emph{observation functions}.

%In this paper, we focus on cooperative MARL scenarios, 
%where agents must collaborate to achieve a common goal.
%In the following, we provide an overview of the key concepts in MARL, 
%including the formalization of the problem, the deep Q-learning algorithm, and the training paradigms used in the literature.
\subsection{Formalization}
Since MARL is a broad field, it is essential to provide a formalization to guide the discussion.
In this paper, we consider partially observable networked markov decision process (PONMDP)~\cite{DBLP:journals/tac/AdlakhaLG12} as a tuple $(\mathcal{G}, \mathcal{S}, \mathcal{A}, \mathcal{O}, \mathcal{P}, \mathcal{R}, \gamma)$, where

@@

%where $\pi = \{\pi^i\}_{i \in N}$ represents the joint policy. 
%The goal in collaborative MARL is often to find a joint policy that maximizes this global average Q-value or a similar global objective.
Several algorithms aim to find the \emph{optimal} Q-value function, which maximizes the expected cumulative reward. 
%
One of the most popular among these is the Deep Q-learning algorithm.
\subsubsection{Deep Q-Learning Formalization}
Deep Q-learning~\cite{DBLP:conf/l4dc/FanWXY20} utilizes a deep neural network, 
often referred to as a Q-network, to approximate the action-value function $Q(s, a)$. 
This network takes the state $s$ as input and outputs a vector of Q-values, one for each possible action $a \in \mathcal{A}$.  
The optimal action-value function, denoted $Q^*(s, a)$, satisfies the Bellman optimality equation:

\[ Q^*(s, a) = \mathbb{E}_{s' \sim \mathcal{P}} \left[ r + \gamma \max_{a'} Q^*(s', a') | s, a \right] \]

where $\mathcal{P}$ is the state transition probability distribution.  
Deep Q-learning aims to train the Q-network to approximate $Q^*(s, a)$ by minimizing the loss function:

@@

where $\theta$ are the parameters of the Q-network, $\theta^-$ are the parameters of a target Q-network (used for stability), 
and $\mathcal{D}$ is the experience replay buffer, a collection of past experiences $(s, a, r, s')$.
%
In the multi-agent setting, each agent $i$ has its own Q-network $Q^i(o^i, a^i; \theta^i)$ that takes its local observation $o^i$ as input.  
For centralized training, the global state $s$ can be used as input. 
The specific form of the multi-agent Q-function and the update rules depend on the chosen MARL algorithm (e.g., distributed Deep Q-learning~\cite{ong2015distributed}, value decomposition methods~\cite{DBLP:conf/icml/SonKKHY19}).  
%For example, in independent Deep Q-learning--more details in \Cref{sec:sota}--each agent updates its Q-network independently using its local reward $r^i$:
%\begin{equation}
%  \resizebox{\linewidth}{!}{$
%  \mathcal{L}^i(\theta^i) = \mathbb{E}_{(o^i, a^i, r^i, o'^i) \sim \mathcal{D}^i} \left[ \left( r^i + \gamma \max_{a'^i} Q^i(o'^i, a'^i; \theta^{i-}) - Q^i(o^i, a^i; \theta^i) \right)^2 \right]
%  $}
%  \end{equation}
%\subsection{Key Concepts}

@@

\paragraph{Centralized Training with Decentralized Execution (CTDE)}
In CTDE, a central learner trains all agents, 
but execution of the learned policy is decentralized. 
This process typically involves two phases:
\begin{enumerate}
  \item \textit{Offline training:} Agents interact with a simulator or environment, gathering experience data used to update the central learner's policy.
  \item \textit{Online execution:} Agents deploy the learned policy in the environment without further learning.
\end{enumerate}
This approach offers several benefits.  
Leveraging a centralized dataset allows for efficient learning, 
potentially exploiting global information, 
which is crucial in cooperative scenarios requiring coordinated actions. 
Furthermore, decentralized execution allows agents to act independently after training.
%
However, CTDE has limitations. 
Scalability can become problematic with a large number of agents. 
The reliance on a central learner hinders adaptation to dynamic environments or tasks during runtime.
\sloppy
\paragraph{Decentralized Training with Decentralized Execution (DTDE)}
DTDE eliminates the central learner; each agent learns independently from its own experiences.
Inter-agent cooperation and communication can be incorporated during both training and execution. 
Unlike CTDE, training and execution phases are not necessarily separate---simultaneous learning and execution allow for adaptation to 
changing environments or tasks at runtime.
%
Key advantages of DTDE include scalability with the number of agents and adaptability to dynamic environments.
%
However, DTDE also presents several challenges to overcome, like: 
i) convergence time is often longer compared to centralized training. 
ii) The lack of central coordination can lead to instability and difficulties in differentiating agents and their environmental interactions; and 
iii) online learning in DTDE can introduce non-stationarity if agents do not account for other agents' evolving policies during training, 
as the environment effectively changes from each agent's perspective.

\subsection{State of the Art}\label{sec:sota}
MARL is a longstanding research area with a rich literature, starting 
from seminal works like~\cite{tan1993multi} and~\cite{busoniu2008comprehensive} to more recent contributions like~\cite{gronauer2022multi} and~\cite{canese2021multi}.
%
In the following, we briefly discuss some of the most relevant approaches in the field, which lead to the motivation for our work.
\subsubsection{Independent Learners}
In independent learners, each agent learns its policy independently, disregarding the actions and policies of other agents. 
This straightforward approach, among the earliest explored in MARL, often relies on variants of Q-learning. 
%
Several algorithms have attempted to address these coordination challenges within the independent learner paradigm. 
Basic decentralized Q-learning~\cite{tan1993multi} allows each agent to learn its Q-values independently but often converges to suboptimal joint policies due to a lack of coordination. 
%Subsequent work introduced heuristic-based extensions to improve coordination.
%Approaches like distributed Q-learning~\cite{lauer2000algorithm} incorporate optimism in Q-value updates and equilibrium selection mechanisms to mitigate the convergence to shadowed equilibria (i.e., equilibria that are unstable due to the independent learning processes). 
%Instead, algorithms such as hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient} and win-or-learn fast policy hill climbing~\cite{bowling2002multiagent} employ adaptive learning rates to address the issue of alternating exploration between agents, improving robustness and convergence.
%
%Empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
Subsequent work introduced heuristic-based extensions to improve coordination, like distributed Q-learning~\cite{lauer2000algorithm}, hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient}, however empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
While initially developed for tabular settings, some of these methods, like hysteretic Q-learning~\cite{palmer2017lenient} and lenient Q-learning, have been extended to deep reinforcement learning settings. 
However, these independent approaches, even with deep learning, often exhibit instability and slow convergence in complex environments and fundamentally lack explicit mechanisms for inter-agent coordination. 
%Consequently, more sophisticated MARL algorithms have been developed that explicitly account for the interactions and policies of other agents to facilitate more effective coordination.
\subsubsection{Neighboring-Based Methods}
Neighboring-based methods in MARL address the scalability and communication overhead challenges of fully centralized approaches by restricting agent interactions to a local neighborhood. 
This introduces agents communication and coordination with a limited set of neighbors, reducing the complexity of full joint action and policy learning.
%
One prominent category within neighboring-based methods is \emph{consensus-based} algorithms. 
These algorithms aim to reach a consensus among agents on shared values or policies through local information exchange. 
Early work like QD-learning~\cite{kar2012qd} extends Q-learning with a consensus update term, 
enabling agents to converge towards optimal Q-values despite only receiving local rewards.
%  
Recent advances in consensus-based MARL leverage deep learning and actor-critic architectures, 
combining local policy updates with consensus updates on critic parameters or gradients, 
as seen in the recent works~\cite{7434032,pennesi2010distributed,zhang2018fully,DBLP:journals/jmlr/CiosekW20}.

@@

\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
This section introduces three neighboring-based distributed learning strategies for multi-agent reinforcement learning.
These strategies leverage local interactions between agents to improve learning efficiency and coordination.
For this investigation, we focus on deep Q-learning as the underlying learning algorithm,
and we consider the following strategies: k-nearest neighbor averaging, k-nearest neighbor consensus, and experience sharing.

\paragraph{K-Nearest Neighbor Definition}
The proposed strategies rely on the concept of k-nearest neighbors.
Formally, consider a \emph{weighted} communication graph $\mathcal{G} = (N, E, W)$, where $N$ is the set of agents, where $W: E \rightarrow \mathbb{R}^+$ is a weight function assigning a positive weight to each edge.  
Let $d(i,j)$ denote the shortest path length between agents $i$ and $j$ in the weighted graph $\mathcal{G}$. 
This can be calculated, for instance, using Dijkstra's algorithm. 
The length of a path is the sum of the weights of the edges along the path. 
If no path exists between $i$ and $j$, we set $d(i,j) = \infty$.
%
Then, the set of $k$-nearest neighbors of agent $i$ is defined as:
$$ \mathcal{N}_k(i) = { j \in N \setminus {i} \mid \text{rank}(d(i,j)) \leq k } $$
where $\text{rank}(d(i,j))$ denotes the rank of the distance $d(i,j)$ among all distances from agent $i$ to other agents in $N \setminus {i}$, sorted in ascending order (e..g, $\text{rank}(d(i,j)) = 1$ for the smallest distance).
Ties can be broken arbitrarily.

\paragraph{K-Nearest Neighbor Averaging}
\sloppy
This approach averages an agent's Q-values with those of its nearest neighbors. 
Each agent maintains a local Q-network and shares its Q-values with its neighbors. 
The neighbors average the received Q-values and use the result to update their own Q-networks. 
This leverages the collective knowledge of neighboring agents to improve learning.
%
Formally, giving $\theta^t_i, i \in N$ the parameters of the Q-networks of the agent $i$ at the time $t$,
the update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \theta^{t+1}_i = \theta^t_i + \alpha \sum_{j \in \mathcal{N}_k(i)} \frac{1}{k} \theta^t_j
\end{equation}

\paragraph{K-Nearest Neighbor Consensus}
In case of homogeneous agents (i.e., when all agents have the same action and observation space),
the goal of learning is to find a common policy that all agents can follow.
This approach is typically used in large-scale systems where agents are identical and need to reach a consensus---see~\cite{DBLP:conf/eusipco/BaldazoPZ19}.
%
In this case, instead of forcing a common policy, we can let agents learn independently and then select the best-performing agent as the reference for the others, leading eventually to a consensus on the best same policy.
%
In this way, each agent $i$ continue to learn independently, but at each iteration, it updates its Q-network using the Q-values of the best-performing agent in its neighborhood.
%
The best performing agent is selected based on the average reward it has received in the last $n$ episodes.
Formally, the update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \theta^{t+1}_i \leftarrow \theta^t_{\text{best}}
\end{equation}
where $\theta^t_{\text{best}}$ is the Q-network of the best-performing agent in the neighborhood $\mathcal{N}_k(i)$ at time $t$.

\paragraph{Experience Sharing}
Experience sharing is a simple yet effective method for distributed learning.
This approach is typically used in centralized training to improve sample efficiency and learning stability~\cite{marl-book}.
In our distributed setting, each agent maintains its own experience replay buffer, 
but it also shares a fraction of its experiences with its neighbors.
This allows agents to learn from each other's experiences, 
potentially accelerating learning and improving coordination because agents can learn from each other's mistakes and successes.
%
Formally, the reply buffer of agent $i$ at time $t$ is $\mathcal{D}^t_i = \{(o^i, a^i, r^i, o'^i)\}$.
At each iteration, agent $i$ shares a fraction $\beta$ of its experiences with its neighbors in $\mathcal{N}_k(i)$.
The update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \mathcal{D}^{t+1}_i = \mathcal{D}^t_i \cup \bigcup_{j \in \mathcal{N}_k(i)} \beta \mathcal{D}^t_j
\end{equation}
\paragraph{Neighboring-Based Training with Deep Q-Learning}
The neighboring-based training strategies can be integrated with deep Q-learning to enable distributed learning in multi-agent settings.
%
The algorithm follow the same structure as the deep Q-learning algorithm, 
but with the addition of the neighboring-based intereactions.
%
This may happen with a different frequency for each strategy, 
depending on the communication overhead and the desired level of coordination.

>>>>

\begin{abstract}
Multi-agent deep reinforcement learning (MADRL) has demonstrated significant potential as a framework for developing autonomous agents capable of operating within complex, multi-agent environments.  
%
Extending the single-agent reinforcement learning framework, MADRL enables the development  of both collaborative and competitive agents a wide range of diverse domains, like robotics, traffic management, and video games.
%
Centralized training with decentralized execution (CTDE) has emerged as the predominant training paradigm, which has demonstrated significant effectiveness in learning complex policies. 
%
However, its reliance on a centralized learner necessitates that agents acquire policies offline and subsequently execute them online. 
This constraint motivates the exploration of decentralized training methodologies.
%
While these techniques offer greater flexibility, they tend to have slower convergence and can suffer from instability compared to centralized approaches.
%
Therefore, this paper investigates the effectiveness of various neighbor-based decentralized training strategies based 
  on the well-known Deep-Q Learning algorithm as a viable alternative to centralized training. 
%  
We evaluate experience sharing, k-nearest neighboring averaging, and k-nearest neighboring consensus methods 
  in a cooperative multi-agent environment and compare their performance against centralized training
  and totally decentralized training. 
%  
Our results show that neighbor-based methods can achieve comparable performance to centralized training 
  while offering improved scalability and communication efficiency. 
%
% Finally, we discuss the trade-offs between these methods and provide insights into their applicability 
%   in different scenarios.
\end{abstract}

@@

\section{Introduction}
The increasing complexity of real-world systems, from robotic swarms to smart grids, demands intelligent control strategies capable of adapting to dynamic and often unpredictable environments.
%
In this direction, \emph{reinforcement learning} (RL)~\cite{sutton2018reinforcement} has emerged as a prominent paradigm for training agents that interact with their environment and learn optimal policies through trial and error.
%
Moving beyond single-agent settings, \emph{multi-agent reinforcement learning} (MARL)~\cite{bucsoniu2010multi} extends RL to scenarios where multiple agents must collaborate or compete to achieve a common goal.
%
Recent advances in deep learning (DL) have further advanced the field, 
enabling the training of complex policies in high-dimensional state and action spaces, 
leading to the emergence of \emph{multi-agent deep reinforcement learning} (MADRL)~\cite{gronauer2022multi}.
%
MADRL has gained significant traction due to its flexibility and capacity to learn diverse behaviors, 
ranging from cooperative~\cite{gupta2017cooperative} to competitive~\cite{tampuu2017multiagent}, 
as demonstrated in various case studies involving video games~\cite{shao2019survey}, robotics~\cite{orr2023multi},
and traffic management~\cite{chu2019multi}. 
%
This paper focuses on the subdomain of \emph{cooperative multi-agent reinforcement learning}, 
a particularly active and impactful area of research within MADRL~\cite{oroojlooy2023review}.

@@

\section{Background and Motivation}\label{sec:background}
Multi-agent reinforcement learning (MARL) involves a group of \emph{agents} interacting with a shared \emph{environment}. 
%
At each timestep, agents receive \emph{observations} and select \emph{actions} according to their individual \emph{policies} (mappings from observations to actions). 
%
These individual actions combine to form a \emph{joint action}, which influences the environment's transition to a new state. 
%
Each agent then receives a scalar \emph{reward} signal, typically a function of the global state and joint action. 
%
The goal of each agent is to maximize its \emph{cumulative long-term reward}. 
%
Environments can be \emph{partially observable}, 
meaning agents may not have access to the full global state, 
only receiving local observations through their respective \emph{observation functions}.

%In this paper, we focus on cooperative MARL scenarios, 
%where agents must collaborate to achieve a common goal.
%In the following, we provide an overview of the key concepts in MARL, 
%including the formalization of the problem, the Deep Q-Learning algorithm, and the training paradigms used in the literature.
\subsection{Formalization}
Since MARL is a broad field, it is essential to provide a formalization to guide the discussion.
In this paper, we consider partially observable networked markov decision process (PONMDP)~\cite{DBLP:journals/tac/AdlakhaLG12} as a tuple $(\mathcal{G}, \mathcal{S}, \mathcal{A}, \mathcal{O}, \mathcal{P}, \mathcal{R}, \gamma)$, where

@@

%where $\pi = \{\pi^i\}_{i \in N}$ represents the joint policy. 
%The goal in collaborative MARL is often to find a joint policy that maximizes this global average Q-value or a similar global objective.
Several algorithms aim to find the \emph{optimal} Q-value function, which maximizes the expected cumulative reward. 
%
One of the most popular among these is the Deep Q-Learning algorithm.
\subsubsection{Deep Q-Learning Formalization}
Deep Q-Learning~\cite{DBLP:conf/l4dc/FanWXY20} utilizes a deep neural network, 
often referred to as a Q-network, to approximate the action-value function $Q(s, a)$. 
This network takes the state $s$ as input and outputs a vector of Q-values, one for each possible action $a \in \mathcal{A}$.  
The optimal action-value function, denoted $Q^*(s, a)$, satisfies the Bellman optimality equation:

\[ Q^*(s, a) = \mathbb{E}_{s' \sim \mathcal{P}} \left[ r + \gamma \max_{a'} Q^*(s', a') | s, a \right] \]
%
where $\mathcal{P}$ is the state transition probability distribution and $r$ is the immediate reward received after taking action $a$ in state $s$.  
Deep Q-Learning aims to train the Q-network to approximate $Q^*(s, a)$ by minimizing the loss function:

@@

where $\theta$ are the parameters of the Q-network, $\theta^-$ are the parameters of a target Q-network (used for stability), 
and $\mathcal{D}$ is the experience replay buffer, a collection of past experiences $(s, a, r, s')$.
%
In the multi-agent setting, each agent $i$ has its own Q-network $Q^i(o^i, a^i; \theta^i)$ that takes its local observation $o^i$ as input.  
For centralized training, the global state $s$ can be used as input. 
The specific form of the multi-agent Q-function and the update rules depend on the chosen MARL algorithm (e.g., distributed Deep Q-Learning~\cite{ong2015distributed}, value decomposition methods~\cite{DBLP:conf/icml/SonKKHY19}).  
%For example, in independent Deep Q-Learning--more details in \Cref{sec:sota}--each agent updates its Q-network independently using its local reward $r^i$:
%\begin{equation}
%  \resizebox{\linewidth}{!}{$
%  \mathcal{L}^i(\theta^i) = \mathbb{E}_{(o^i, a^i, r^i, o'^i) \sim \mathcal{D}^i} \left[ \left( r^i + \gamma \max_{a'^i} Q^i(o'^i, a'^i; \theta^{i-}) - Q^i(o^i, a^i; \theta^i) \right)^2 \right]
%  $}
%  \end{equation}
%\subsection{Key Concepts}

@@

\paragraph{Centralized Training with Decentralized Execution (CTDE)}
In CTDE, a central learner trains all agents, 
but execution of the learned policy is decentralized. 
This process typically involves two phases:
\begin{enumerate}
  \item \textit{Offline training:} agents interact with a simulator or environment, gathering experience data used to update the central learner's policy.
  \item \textit{Online execution:} agents deploy the learned policy in the environment without further learning.
\end{enumerate}
This approach offers several benefits.  
Leveraging a centralized dataset allows for efficient learning, 
potentially exploiting global information, 
which is crucial in cooperative scenarios requiring coordinated actions. 
Furthermore, decentralized execution allows agents to act independently after training.
%
However, CTDE has limitations. 
Scalability can become problematic with a large number of agents. 
The reliance on a central learner hinders adaptation to dynamic environments or tasks during runtime.
\sloppy
\paragraph{Decentralized Training with Decentralized Execution (DTDE)}
DTDE eliminates the central learner; each agent learns independently from its own experiences.
Inter-agent cooperation and communication can be incorporated during both training and execution. 
Unlike CTDE, training and execution phases are not necessarily separate---simultaneous learning and execution allow for adaptation to 
change environments or tasks at runtime.
%
Key advantages of DTDE include scalability with the number of agents and adaptability to dynamic environments.
%
However, DTDE also presents several challenges to overcome, like: 
i) convergence time is often longer compared to centralized training; 
ii) the lack of central coordination can lead to instability and difficulties in differentiating agents and their environmental interactions; and 
iii) online learning in DTDE can introduce non-stationarity if agents do not account for other agents' evolving policies during training, 
as the environment effectively changes from each agent's perspective.

\subsection{State of the Art}\label{sec:sota}
MARL is a longstanding research area with a rich literature, starting 
from seminal works like~\cite{tan1993multi} and~\cite{busoniu2008comprehensive} to more recent contributions like~\cite{gronauer2022multi} and~\cite{canese2021multi}.
%
In the following, we briefly discuss some of the most relevant approaches in the field, which lead to the motivation for our work.
\subsubsection{Independent Learners}
In independent learners, each agent learns its policy independently, disregarding the actions and policies of other agents. 
This straightforward approach, among the earliest explored in MARL, often relies on variants of Q-learning. 
%
Several algorithms have attempted to address these coordination challenges within the independent learner paradigm. 
Basic decentralized Q-learning~\cite{tan1993multi} allows each agent to learn its Q-values independently but often converges to suboptimal joint policies due to a lack of coordination. 
%Subsequent work introduced heuristic-based extensions to improve coordination.
%Approaches like distributed Q-learning~\cite{lauer2000algorithm} incorporate optimism in Q-value updates and equilibrium selection mechanisms to mitigate the convergence to shadowed equilibria (i.e., equilibria that are unstable due to the independent learning processes). 
%Instead, algorithms such as hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient} and win-or-learn fast policy hill climbing~\cite{bowling2002multiagent} employ adaptive learning rates to address the issue of alternating exploration between agents, improving robustness and convergence.
%
%Empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
Subsequent work introduced heuristic-based extensions to improve coordination, like distributed Q-learning~\cite{lauer2000algorithm}, hysteretic Q-learning~\cite{matignon2007hysteretic}, lenient Q-Learning~\cite{bloembergen2010lenient}, however empirical comparisons of these algorithms highlight the trade-offs in performance across various MARL tasks and coordination challenges~\cite{matignon2012independent}. 
While initially developed for tabular settings, some of these methods, like hysteretic Q-learning~\cite{palmer2017lenient},
%and lenient Q-learning, 
have been extended to deep reinforcement learning settings. 
However, these independent approaches, even with deep learning, often exhibit instability and slow convergence in complex environments and fundamentally lack explicit mechanisms for inter-agent coordination. 
%Consequently, more sophisticated MARL algorithms have been developed that explicitly account for the interactions and policies of other agents to facilitate more effective coordination.
\subsubsection{Neighboring-Based Methods}
Neighboring-based methods in MARL address the scalability and communication overhead challenges of fully centralized approaches by restricting agent interactions to a local neighborhood. 
This introduces agents communication and coordination with a limited set of neighbors, reducing the complexity of full joint action and policy learning.
%
One prominent category within neighboring-based methods is \emph{consensus-based} algorithms. 
These algorithms aim to reach a consensus among agents on shared values or policies through local information exchange. 
Early work like QD-learning~\cite{kar2012qd} extends Q-learning with a consensus update term, 
enabling agents to converge towards optimal Q-values despite only receiving local rewards.
%  
Recent advances in consensus-based MARL leverage deep learning and actor-critic architectures, 
combining local policy updates with consensus updates on critic parameters or gradients, 
as seen in the recent works~\cite{7434032,pennesi2010distributed,zhang2018fully,DBLP:journals/jmlr/CiosekW20}.

@@

\section{Neighboring-Based Distributed Learning Strategies}\label{sec:neighboring}
This section introduces three neighboring-based distributed learning strategies for multi-agent reinforcement learning.
These strategies leverage local interactions between agents to improve learning efficiency and coordination.
For this investigation, we focus on Deep Q-Learning as the underlying learning algorithm,
and we consider the following strategies: k-nearest neighbor averaging, k-nearest neighbor consensus, and experience sharing.

\paragraph{K-Nearest Neighbor Definition}
The proposed strategies rely on the concept of k-nearest neighbors.
Formally, consider a \emph{weighted} communication graph $\mathcal{G} = (N, E, W)$, where $N$ is the set of agents, where $W: E \rightarrow \mathbb{R}^+$ is a weight function assigning a positive weight to each edge.  
Let $d(i,j)$ denote the shortest path length between agents $i$ and $j$ in the weighted graph $\mathcal{G}$. 
This can be calculated, for instance, using Dijkstra's algorithm. 
The length of a path is the sum of the weights of the edges along the path. 
If no path exists between $i$ and $j$, we set $d(i,j) = \infty$.
%
Then, the set of $k$-nearest neighbors of agent $i$ is defined as:
$$ \mathcal{N}_k(i) = { j \in N \setminus {i} \mid \text{rank}(d(i,j)) \leq k } $$
where $\text{rank}(d(i,j))$ denotes the rank of the distance $d(i,j)$ among all distances from agent $i$ to other agents in $N \setminus {i}$, sorted in ascending order (e.g, $\text{rank}(d(i,j)) = 1$ for the smallest distance).
Ties can be broken arbitrarily.

\paragraph{K-Nearest Neighbor Averaging}
\sloppy
%This approach averages an agent's Q-values with those of its nearest neighbors. 
In this strategy, each agent maintains a local Q-network and shares its Q-values with its neighbors.
The neighbors average the received Q-values and use the result to update their own Q-networks. 
This approach allows agents to learn from their neighbors' experiences, potentially improving learning efficiency and coordination.
%
Formally, giving $\theta^t_i, i \in N$ the parameters of the Q-networks of the agent $i$ at the time $t$,
the update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \theta^{t+1}_i = \theta^t_i + \alpha \sum_{j \in \mathcal{N}_k(i)} \frac{1}{k} \theta^t_j
\end{equation}

\paragraph{K-Nearest Neighbor Consensus}
In case of homogeneous agents (i.e., when all agents have the same action and observation space),
the goal of learning is to find a common policy that all agents can follow.
This approach is typically used in large-scale systems where agents are identical and need to reach a consensus---see~\cite{DBLP:conf/eusipco/BaldazoPZ19}.
%
Instead of directly enforcing a shared policy, 
we allow each agent to learn independently and periodically select the best-performing agent within its neighborhood as a reference. 
This approach facilitates decentralized exploration while promoting consensus towards a high-performing policy.
%
Formally, Let $R_i(t)^n$ be the average reward obtained by agent $i$ over the last $n$ episodes, namely
$R_i(t)^n = \frac{1}{n} \sum_{k=t-n}^{t} r_i(k)$.
At each time step $t$, 
each agent $i$ identifies the best-performing neighbor $b_i(t)$ within its $k$-neighborhood $\mathcal{N}_i^k$:

$$b_i(t) = \underset{j \in \mathcal{N}_k(i)}{\operatorname{argmax}} R_j(t)^n$$
%
Agent $i$ then updates its Q-network parameters $\theta_i$ using the parameters of the best-performing neighbor, namely:
$\theta_i^{t+1} \leftarrow \theta_{b_i(t)}^t$.
%
This process repeats at every consensus update interval.  
Choosing an appropriate interval balances independent exploration with consensus convergence.

\paragraph{Experience Sharing}
Experience sharing is a simple yet effective method for distributed learning.
This approach is typically used in centralized training to improve sample efficiency and learning stability~\cite{marl-book}.
In our distributed setting, each agent maintains its own experience replay buffer, 
but it also shares a fraction of its experiences with its neighbors.
%
This allows agents to learn from each other's experiences, 
potentially accelerating learning and improving coordination because agents can learn from each other's mistakes and successes.
%
Formally, the reply buffer of agent $i$ at time $t$ is $\mathcal{D}^t_i = \{(o^i, a^i, r^i, o'^i)\}$.
At each iteration, agent $i$ shares a fraction $\beta$ of its experiences with its neighbors in $\mathcal{N}_k(i)$.
The update rule for agent $i$ at time $t+1$ is:
\begin{equation}
  \mathcal{D}^{t+1}_i = \mathcal{D}^t_i \cup \bigcup_{j \in \mathcal{N}_k(i)} \beta \mathcal{D}^t_j
\end{equation}
\paragraph{Neighboring-Based Training with Deep Q-Learning}
\sloppy
The neighboring-based training strategies can be integrated with Deep Q-Learning to enable distributed learning in multi-agent settings.
%
The algorithm follow the same structure as the Deep Q-Learning algorithm, 
but with the addition of the neighboring-based interactions.
%
This may happen with a different frequency for each strategy, 
depending on the communication overhead and the desired level of coordination.
%
In \Cref{alg:neighbor_dqn} we provide a detailed description of the integration of the neighboring-based strategies with Deep Q-Learning.
\begin{algorithm}
  \caption{Neighboring-Based Deep Q-Learning}
  \label{alg:neighbor_dqn}
  \ForEach{agent $i \in N$}{
      Initialize Q-network $Q^i(o^i, a^i; \theta^i)$ and target network $Q^{i-}(o^i, a^i; \theta^{i-})$\\
      Initialize replay buffer $\mathcal{D}^i$
  }
  \For{episode = 1 to M}{
      \ForEach{agent $i \in N$}{
          Initialize initial observation $o_0^i$
      }
      \For{t = 0 to T}{
          \ForEach{agent $i \in N$}{
              Select action $a_t^i$ based on $\epsilon$-greedy policy using $Q^i(o_t^i, a^i; \theta^i)$\\
              Execute action $a_t^i$ and observe reward $r_{t+1}^i$ and next observation $o_{t+1}^i$\\
              Store $(o_t^i, a_t^i, r_{t+1}^i, o_{t+1}^i)$ in $\mathcal{D}^i$\\
              Sample a mini-batch from $\mathcal{D}^i$\\
              Calculate target values using chosen neighboring strategy (see below)\\
              Update $\theta^i$ by minimizing the TD error\\
              Every C steps, update target network $\theta^{i-} \leftarrow \theta^i$\\
              \textbf{
              Every S steps, perform neighboring strategy}
          }
      }
  }
\end{algorithm}
\section{Evaluation}\label{sec:experiments}
This section presents the experimental setup and evaluation metrics used to assess the performance of the proposed neighboring-based training strategies.
%
All the experiments and data produced are reproducible and publicly available\footnote{\url{https://anonymous.4open.science/r/experiments-2024-SAC-marl-neighbouring-strategies-4B8F}}.
\subsection{Simulated environment}
The proposed multi-agent environment is a variant of the ``gather the items'' scenario, 
designed to study the impact of neighborhood strategies. 
It consists of a 2D Euclidean space populated by $A$ agents and $T$ targets, 
initially placed at random locations.
The objective is for the agents to work together to capture all targets, resulting in a successful termination of the environment episode.
A maximum time limit of $T_{max} = 300$ time steps is enforced to prevent indefinite episodes.
